\chapter*{AUTHORSHIP STATEMENT}

This dissertation was authored by the student, ATM Jahid Hasan,
in partial fulfillment of the requirements of the Doctor of Philosophy degree.
The work presented here includes chapters that were developed
in collaboration with other researchers.

Chapter \ref{chap:diff} describes research primarily performed
at North Carolina State University.
The student performed MD simulations, analyzed the data,
and prepared the manuscript detailing the study.
Dr. Benjamin Beeler supervised the study
and reviewed and edited the manuscript.
The following manuscript has been published:
ATM Jahid Hasan and Benjamin Beeler.
Calculation of grain boundary diffusion coefficients in $\gamma$U-Mo
using atomistic simulations.
Journal of Nuclear Materials, page 155190, 2024.

Chapter \ref{chap:xeres} includes research performed at
Idaho National Laboratory and North Carolina State University.
The student devised the research plan,
programmed the software necessary for the research,
analyzed simulation data, and prepared the manuscript.
Dr. Linu Malakkal and Dr. Mathew Swisher from Idaho National Laboratory
assisted with the MD simulations and reviewed the manuscript.
Dr. Benjamin Beeler supervised the entire research process
and reviewed and edited the manuscript.

Chapter \ref{chap:iuq} details research conducted at
Argonne National Laboratory and North Carolina State University.
The data for the research was supplied by
Dr. Bei Ye, Dr. Zhi-Gang Mei, and Dr. Park Gyuchul
from Argonne National Laboratory.
The student performed the analysis and prepared the manuscript.
Dr. Benjamin Beeler supervised the research
and reviewed and edited the manuscript.

\clearpage
