\chapter*{DEDICATION}

\begin{center}
	% TODO: make sure the spelling is right
	To my late grandmother, \textit{Johra Begum}.

	It has been a long time,
	but I can still feel her fingers through my hair as I fall asleep.
\end{center}

\clearpage

\chapter*{BIOGRAPHY}

ATM Jahid Hasan was born in Dhaka, Bangladesh.
He also spent the first 25 years of his life there.
ATM Jahid received his Bachelor of Science degree in Nuclear Engineering
in 2019 from the University of Dhaka.
In 2021, he moved to the US to attend graduate school.
So far, he has received a Graduate Cerificate in Materials Informatics
and a Master of Nuclear Engineering degree
from the North Carolina State University.
His research has focused on atomistic simulations
and other computational methods.
He likes to solve problems and contributes to open-source software.

\clearpage

\chapter*{ACKNOWLEDGEMENTS}

I would like to thank Dr. Benjamin Beeler for his supervision.
He has guided my research and academic growth with great care.
He has also been extremely patient with me over the years.
I have been lucky to have him as my supervisor.
I would also like to express my gratitude to my committee members,
Dr. Jacob Eapen, Dr. Xu Wu, and Dr. Jun Liu, for their time and insight.
I have learned a great deal from Dr. Eapen and Dr. Wu,
both through our conversations and the courses they taught.

My gratitude also extends to my mentors from the national laboratories.
Dr. Linu Malakkal hosted me at Idaho National Laboratory for two summers
with great enthusiasm.
Dr. Bei Ye and Dr. Zhi-Gang Mei have guided me
throughout my time at Argonne National Laboratory.
They have continued to be my collaborators
and supported my research tremendously.

Thanks to my colleagues and friends at the Nuclear Engineering department:
Shehab Shousha, Kai Duemmler, Khadija Mahbuba, Syed Rizvi, Matt deJong,
Philip Alarcon-Furman, Fu-Yun Tsai, Da Cao, Saqr Alshogeathri, Logan Williams
and others.
A special thanks to Mario Milev, who has always been there for any kind of help.
Also, thanks to Professor Lisa Marshall
for keeping the department lively all the time.

I am grateful to my parents, Anamul and Jakia,
and I appreciate their perseverance.
They have dedicated all their effort and resources to make me succeed.
It definitely has been difficult for them
not to be able to see me years at a time.
Also, shoutout to my brother, Opar,
who now has to deal with my parents on behalf of both of us every day.
I would also like to thank my in-laws: Luis, Katy, Ale and Lia.
They have welcomed me into their family with a lot of love.

I must also thank Mishti,
who thinks it is her right to jump on to my lap and make biscuits
whenever I sit down to work.
I suspect I am doing all this work to provide her with treats,
and I am extremely glad that she came into my life
with tiny paws and big demands.

Finally, I would thank my wife, Lucia Rebeca Gomez Hurtado.
She is a wonderful person with a kind heart and an ocean of empathy.
She has been my refuge throughout these last few years.
This chapter of my life has been a privilege,
and I am grateful to everyone who helped me along the way.

\clearpage
